---
title: "Hora-1 Raw Material Survey"
author: "Caleb Chen"
date: "5/24/2022"
output:
  html_document: default
  pdf_document: default
---
Loading packages
```{r}
library(tidyverse)
library(tidyr)
library(spatstat)
library(here)
library(readr)
library(dplyr)
library(insight)
library(gridExtra)
library(lattice)
library(viridis)
library(ggsci)
library(reactable)
library(shadowtext)
library(ggrepel)
```
#1.0 Loading flake data and calculating surface area and cortical surface area
```{r}
final_flake_data = read.csv("C:/Users/caleb/Desktop/Hora-1 R Code and Data/hor-1_cortex_ratio.csv")
raw_material_survey = read.csv("C:/Users/caleb/Desktop/Hora-1 R Code and Data/raw_material_survey_6.csv")

#calculate cortical surface area in cm^2. Note that I recorded flake dimensions in mm, so I convert those to cm^2)
surface_area = ((final_flake_data$MaxLength/10)*(final_flake_data$MaxWidth/10))
cortical_sa = ((final_flake_data$cortex_mp*(final_flake_data$MaxLength/10)*(final_flake_data$MaxWidth/10)))

#calculate total assemblage volume. Volume = mass/quartz density (2.64 from Berman method)
artifact_volume = ((final_flake_data$Mass/2.64))

#pipe the new cortical_sa and volume columns into the df
final_flake_data_new <- final_flake_data %>%
  mutate(surface_area, cortical_sa, artifact_volume)
```

##1.1 Hora 1 flake statistics
```{r}
max(final_flake_data_new$MaxLength[!is.na(final_flake_data_new$MaxLength)])
min(final_flake_data_new$MaxLength[!is.na(final_flake_data_new$MaxLength)])
mean(final_flake_data_new$MaxLength[!is.na(final_flake_data_new$MaxLength)])
median(final_flake_data_new$MaxLength[!is.na(final_flake_data_new$MaxLength)])
sd(final_flake_data_new$MaxLength[!is.na(final_flake_data_new$MaxLength)])
```

# 2.0 Calculating Average Cortical Surface Area for all Survey Nodules
```{r average raw material cortical surface area for all survey nodules}
#average cortical surface area of modern raw material nodule (in cm^2)
average_raw_mat_nodule_cortical_sa = mean(raw_material_survey$cortical_surface_area[!is.na(raw_material_survey$cortical_surface_area)])
round(average_raw_mat_nodule_cortical_sa, 2)
```

# 3.0 Attribute Counts for All Survey Nodules
## 3.1 Raw Material Shape Counts
```{r raw material shape counts}
ggplot(raw_material_survey)+
  geom_histogram(mapping = aes(x=shape), stat="count", fill="turquoise") +
  labs(x="Shape", y="Count")+
  theme(text = element_text(size = 14))

ggsave("raw_mat_shape.png")
```

## 3.2 Raw Material Texture Counts
```{r raw material texture counts}
ggplot(raw_material_survey) +
  geom_histogram(mapping = aes(factor(texture, levels=c("coarse","medium","fine"))), stat="count", fill="salmon") +
  labs(x="Texture", y="Count")+
  theme(text = element_text(size = 14))

ggsave("raw_mat_texture.png")
```
## 3.3 Raw Material Transparency Counts
```{r raw material transparency counts}
ggplot(raw_material_survey)+
  geom_histogram(mapping = aes(x=transparency), stat="count", fill="turquoise") +
  labs(x="transparency", y="Count")+
  theme(text = element_text(size = 14))

ggsave("raw_mat_transparency.png")
```

# 4.0 Cortex Percentages
```{r raw material nodule cortex percentages}
ggplot(raw_material_survey) +
  geom_bar(mapping = aes((x=factor(cortex, levels=c("0","1-9","10-39", "40-59","60-89", "90-99", "100")))), stat="count", fill="salmon") +
  labs(x="Cortex Percentage", y="Count")+
  theme(text = element_text(size = 14)) +
  ggtitle("Survey Raw Material Sample Cortex Percentages")
table(raw_material_survey$cortex)
nodule_table= prop.table(table(raw_material_survey$cortex))
round(nodule_table, 2)

ggsave("raw_mat_cortex_total.png")

raw
```
# 5.0 Survey Nodule Attribute Counts per Location
## 5.1 Survey Nodule Cortex Percentage per Location
```{r raw material cortex percentage per location}

mean(raw_material_survey$cortex_midpoint)
median(raw_material_survey$cortex_midpoint)
sd(raw_material_survey$cortex_midpoint)

raw_material_survey %>%
  group_by(location) %>%
  count(cortex) %>%
  ggplot(aes(x = as.factor(location), y = n, fill = factor(cortex, levels=c("0","1-9","10-39", "40-59","60-89", "90-99", "100")))) +
  geom_col(position = "fill",colour="black")+
  scale_fill_jco() +
  labs(x="location",y="proportion", fill='cortex percentage')+
  geom_text_repel(aes(label = paste(n)), 
                lineheight = 2,
            fontface = "bold", bg.colour = "white",bg.r = .1, force = 0,
                position = position_fill(vjust = 0.5), size=3)

ggsave("raw_mat_cortex_facetted.png")
```
## 5.2 Survey Nodule Structure Counts per Location
```{r raw material structure per location}
raw_material_survey %>%
  group_by(location) %>%
  count(structure) %>%
  ggplot(aes(x = as.factor(location), y = n, fill = structure)) +
  geom_col(position = "fill", colour="black")+
  scale_fill_jco() +
  labs(x="location",y="proportion")+
  geom_text_repel(aes(label = paste(n)), 
                lineheight = 2,
            fontface = "bold", bg.colour = "white",bg.r = .1, force = 0,
                position = position_fill(vjust = 0.5), size=3)

ggsave("raw_mat_structure_facetted.png")
```
## 5.3 Survey Nodule Shape Counts per Location
```{r raw material shape per location}
raw_material_survey %>%
  group_by(location) %>%
  count(shape) %>%
  ggplot(aes(x = as.factor(location), y = n, fill = shape)) +
  geom_col(position = "fill",colour="black")+
  scale_fill_jco() +
  labs(x="location",y="proportion")+
  geom_text_repel(aes(label = paste(n)), 
                lineheight = 2,
            fontface = "bold", bg.colour = "white",bg.r = .1, force = 0,
                position = position_fill(vjust = 0.5), size=3)

ggsave("raw_mat_shape_facetted.png")
```
## 5.4 Survey Nodule Texture Counts per Location
```{r raw material texture per location}
raw_material_survey %>%
  group_by(location) %>%
  count(texture) %>%
  ggplot(aes(x = as.factor(location), y = n, fill = texture)) +
  geom_col(position = "fill", colour="black")+
  labs(x="location",y="proportion")+
  scale_fill_jco() +
  geom_text_repel(aes(label = paste(n)), 
                lineheight = 2,
            fontface = "bold", bg.colour = "white",bg.r = .1, force = 0,
                position = position_fill(vjust = 0.5), size=3)

ggsave("raw_mat_texture_facetted.png")
```
## 5.5 Survey Nodule Transparency Counts per Location
```{r transparency per location}
raw_material_survey %>%
  group_by(location) %>%
  count(transparency) %>%
  ggplot(aes(x = as.factor(location), y = n, fill = transparency)) +
  geom_col(position = "fill", colour="black")+
  scale_fill_jco() +
  labs(x="location",y="proportion")+
  geom_text_repel(aes(label = paste(n)), 
                lineheight = 2,
            fontface = "bold", bg.colour = "white",bg.r = .1, force = 0,
                position = position_fill(vjust = 0.5), size=3)

ggsave("raw_mat_transparency.png")
```

## 5.6 Total Survey Nodule Surface Area per Location

```{r Total raw material surface area per location}
ggplot(raw_material_survey, aes(x= as.factor(location), y = surface_area)) +
  geom_boxplot() +
  labs(x="location", y="surface area (cm^2)")+
  scale_fill_jco() 

ggsave("raw_mat_surface_area.png")
```
##5.7 Survey nodule Surface Area per Location (separated by raw material)
```{r raw material surface area per location separated by raw material}
ggplot(raw_material_survey, aes(x= as.factor(location), y = surface_area)) +
  geom_boxplot() +
  facet_wrap(vars(transparency))+
  labs(x="location", y="surface area (cm^2)")+
  scale_fill_jco() 

ggsave("raw_mat_surface_area_facetted.png")
```
## 5.8 Survey Nodule Cortical Surface Area per Location
```{r raw material cortical surface area per location}
#I will fix the boxplot axes in the code 
ggplot(raw_material_survey, aes(x= as.factor(location), y = cortical_surface_area)) +
  geom_boxplot() +
  facet_wrap(vars(transparency))+
  labs(x="location", y="cortical surface area (cm^2)")

raw_mat_csa_anova = aov(cortical_surface_area ~ as.factor(location), data=raw_material_survey)
summary(raw_mat_csa_anova)

ggsave("raw_mat_csa_per_location.png")
```


# 6.0 Survey Nodule Statistics
## 6.1 Survey Nodule Cortex Midpoint Statistics
```{r cortex midpoint value statistics}
mean(raw_material_survey$cortex_midpoint[!is.na(raw_material_survey$cortex_midpoint)])
sd(raw_material_survey$cortex_midpoint[!is.na(raw_material_survey$cortex_midpoint)])
median(raw_material_survey$cortex_midpoint[!is.na(raw_material_survey$cortex_midpoint)])
```
## 6.2 Survey Nodule Cortical Surface Area Statistics
```{r survey nodule cortical surface area statistics}
mean(raw_material_survey$cortical_surface_area[!is.na(raw_material_survey$cortical_surface_area)])
sd(raw_material_survey$cortical_surface_area[!is.na(raw_material_survey$cortical_surface_area)])
median(raw_material_survey$cortical_surface_area[!is.na(raw_material_survey$cortical_surface_area)])
```

# 7.0 Hora-1 Flake Statistics
## 7.1 Hora-1 Flake Cortical Surface Area Statistics
```{r Checking data normality for plotted and sieved crystal and vein quartz cortical surface areas}
ggplot(final_flake_data_new, aes(y = cortical_sa)) +
  geom_histogram() +
  facet_grid(vars(SievedPlotted, Age, RawMaterial))
```


# 8.0 Cortex Ratio Calculations
## 8.1 Crystal and Vein Quartz Cortex Ratio Calculations
# Cortex Ratio Calculations
```{r Cortex Ratio Calculations}
TP_crystal <- subset(final_flake_data_new, RawMaterial == 'CrystalQuartz'| Age == 'TerminalPleistocene')
TP_vein <- subset(final_flake_data_new, RawMaterial == 'VeinQuartz'| Age == 'TerminalPleistocene')
TLGM_crystal <- subset(final_flake_data_new, RawMaterial == 'CrystalQuartz'| Age == 'TerminalLGM')
TLGM_vein <- subset(final_flake_data_new, RawMaterial == 'VeinQuartz'| Age == 'TerminalLGM')
  
TP_crystal_obs = sum(TP_crystal$cortical_sa[!is.na(TP_crystal$cortical_sa)])
TP_vein_obs = sum(TP_vein$cortical_sa[!is.na(TP_vein$cortical_sa)])
TLGM_crystal_obs = sum(TLGM_crystal$cortical_sa[!is.na(TLGM_crystal$cortical_sa)])
TLGM_vein_obs = sum(TLGM_vein$cortical_sa[!is.na(TLGM_vein$cortical_sa)])

#expected cortical surface area for Terminal Pleistocene nodules
crystal_nodules = subset(raw_material_survey, transparency == 'translucent')
no_TP_crystal_nodules = (sum(TP_crystal$artifact_volume[!is.na(TP_crystal$artifact_volume)]))/(mean(crystal_nodules$volume[!is.na(crystal_nodules$volume)]))
no_TLGM_crystal_nodules =(sum(TLGM_crystal$artifact_volume[!is.na(TLGM_crystal$artifact_volume)]))/(mean(crystal_nodules$volume[!is.na(crystal_nodules$volume)]))
TP_crystal_exp = mean(crystal_nodules$cortical_surface_area[!is.na(crystal_nodules$cortical_surface_area)])*no_TP_crystal_nodules
TLGM_crystal_exp = mean(crystal_nodules$cortical_surface_area[!is.na(crystal_nodules$cortical_surface_area)])*no_TLGM_crystal_nodules

#expected cortical surface area for vein quartz
vein_nodules = subset(raw_material_survey, transparency == 'opaque')
no_TP_vein_nodules = (sum(TP_vein$artifact_volume[!is.na(TP_vein$artifact_volume)]))/(mean(vein_nodules$volume[!is.na(vein_nodules$volume)]))
no_TLGM_vein_nodules = (sum(TLGM_vein$artifact_volume[!is.na(TLGM_vein$artifact_volume)]))/(mean(vein_nodules$volume[!is.na(vein_nodules$volume)]))
TP_vein_exp = mean(vein_nodules$cortical_surface_area[!is.na(vein_nodules$cortical_surface_area)])*no_TP_vein_nodules
TLGM_vein_exp = mean(vein_nodules$cortical_surface_area[!is.na(vein_nodules$cortical_surface_area)])*no_TLGM_vein_nodules

#cortex ratios
TP_crystal_ratio = TP_crystal_obs/TP_crystal_exp
TP_vein_ratio = TP_vein_obs/TP_vein_exp
TLGM_crystal_ratio = TLGM_crystal_obs/TLGM_crystal_exp
TLGM_vein_ratio = TLGM_vein_obs/TLGM_vein_exp

TP_crystal_ratio
TP_vein_ratio
TLGM_crystal_ratio
TLGM_vein_ratio
```
##8.2 ANOVA tests for cortical surface area differences between plotted and sieved lithics
```{r Hora-1 flake cortical surface area statistics}
mean(final_flake_data_new$cortical_sa[!is.na(final_flake_data_new$cortical_sa)])
sd(final_flake_data_new$cortical_sa[!is.na(final_flake_data_new$cortical_sa)])
median(final_flake_data_new$cortical_sa[!is.na(final_flake_data_new$cortical_sa)])

TLGM_cq_csa_anova = aov(cortical_sa ~ SievedPlotted , data=TLGM_crystal)
TLGM_vein_csa_anova = aov(cortical_sa ~ SievedPlotted , data=TLGM_vein)
summary(TLGM_cq_csa_anova)
summary(TLGM_vein_csa_anova)

ggplot(final_flake_data_new, aes(x= SievedPlotted, y = log(cortical_sa), fill=RawMaterial)) +
  geom_boxplot() +
  facet_grid(vars(Age, RawMaterial))+
  labs(x="sieved/plotted", y="cortical surface area (cm^2)")+
  theme(text = element_text(size = 9))+
  scale_fill_jco()

ggsave("HOR-1_flake_csa_plotted_sieved.png")
```

# 9.0 Cortical Surface Area Plots
## 9.1 Survey Nodule Cortical Surface Areas per Nodule Attribute
```{r}
ggplot(raw_material_survey) +
  geom_boxplot(mapping = aes(x=shape, y=cortical_surface_area)) +
  labs(x="shape", y="cortical surface area (cm^2)")+
  theme(text = element_text(size = 14))

ggsave("cortical_sa.png")

ggplot(raw_material_survey) +
  geom_boxplot(mapping = aes(x=transparency, y=cortical_surface_area)) +
  labs(x="transparency", y="cortical surface area (cm^2)")+
  theme(text = element_text(size = 11))

ggplot(raw_material_survey) +
  geom_boxplot(mapping = aes(x=transparency, y=surface_area)) + 
  labs(x="transparency", y="surface area (cm^2)")+
  theme(text = element_text(size=12))

ggplot(raw_material_survey) +
  geom_boxplot(mapping = aes(x=transparency, y=volume)) + 
  labs(x="transparency", y="volume (cm^3)")+
  theme(text = element_text(size=12))
```


# 13.0 Survey and Artifact Comparison
```{r transparency raw material and artifact}
crystallinity = read.csv("C:/Users/caleb/Desktop/MALAPP_raw_material_survey (1)/crystallinity_2.csv")
ggplot(crystallinity, aes(x=source, y=count, fill=material)) +
  theme(text = element_text(size = 16)) +
  geom_bar(position='fill', stat='identity',colour="black")+
  scale_fill_jco() +
  labs(x="source", y="proportion") +
  geom_text_repel(aes(label = paste(count)), 
                lineheight = 2,
            fontface = "bold", bg.colour = "white",bg.r = .1, force = 0,
                position = position_fill(vjust = 0.5), size=5)
  

ggsave("artifact_vs_survey.png")
```



```{r}
sample = read.csv("C:/Users/caleb/Desktop/MALAPP_raw_material_survey (1)/sample.csv")
sample_n(sample, 20)
```

